---
title: "Modelling overdispersion"
author: "Mangiola Stefano"
date: "15/06/2018"
output:
  html_document:
    toc: yes
    toc_depth: 3
    code_folding: hide
  github_document:
    toc: yes
    toc_depth: 3
linkcolor: magenta
urlcolor: magenta
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, results='hide', include=FALSE}
library(foreach)
library(tidyverse)
library(doParallel)
registerDoParallel()
```

Here I show at what extent we can approximate overdispersed poisson with negative binomial or gamma distributions

## Both negative binomial and gamma can limit to a poisson

```{r limit_to_poisson}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %do% {
	s=1

	x = rgamma(10000,shape = m * s, rate = s)
	
	c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% 
	rename("gamma" = `variance.10%`) %>% 
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1000000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% 
			rename("nb" = `variance.10%`),
	by="mean"
	) %>%
		
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
		gather(label, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=label)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```


However gamma and negative binomial introduce overdispersion "generalising" the relation between expected value and variance of the poisson in a different way

## Gamma with increasing overdispersion - fixing rate  
**rgamma(10000,shape = m * s, rate = s)**  

```{r gamma}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		x = rgamma(10000,shape = m * s, rate = s)
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% 
	rename("1" = `variance.10%`) %>% 
		
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.8
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("0.8" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.6
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("0.6" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.4
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("0.4" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.2
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("0.2" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.1
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("0.1" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
		gather(rate_param_R_gamma, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=rate_param_R_gamma)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

## Negative binomial with increasing overdispersion  
**rnbinom(10000, mu = m, size = s)**  

```{r negative_binomial}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Negative binomial
	foreach(m = mm, .combine = bind_rows) %do% {
			s=50000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% 
	rename("50000" = `variance.10%`) %>% 
		
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=20000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("20000" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=10000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("10000" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=5000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("5000" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("1000" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=500
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("500" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
	
		gather(size_param_R_nb, variance, -mean) 

```

```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=size_param_R_nb)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

## Log normal with increasing SD  
**exp(rnorm(10000, log(m), log(s)))**  

```{r log normal}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Negative binomial
	foreach(m = mm, .combine = bind_rows) %do% {
			s=1.02
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% 
	rename("1.02" = `variance.10%`) %>% 
		
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.03
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("1.03" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.04
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("1.04" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.05
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("1.05" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.06
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("1.06" = `variance.10%`),
		by="mean"
	) %>%
	
	# Negative binomial
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.07
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("1.07" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
		gather(sd, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=sd)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

## poisson + Gamma + negative binomial + log-normal

Gamma and negative binomial reach same level of overdispersion with two different trajectories

```{r gamma and negative_binomial}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Negative binomial
	foreach(m = mm, .combine = bind_rows) %do% {
			s=3000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("negative binomial" = `variance.10%`) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.4
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("gamma" = `variance.10%`),
		by="mean"
	) %>%
	
	# Log-normal
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=1.025
			x = exp(rnorm(10000, log(m), log(s)))
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("log-normal" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
	gather(Label, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=Label)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```


```{r gamma reparametrised, echo=FALSE, eval=FALSE}

## Reparametrise the gamma 

#to have constant SD and apply a linear dependence between SD and mean

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.3 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% 
	rename("0.3" = `variance.10%`) %>% 
		
	# Gamma
	left_join(
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.25 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% rename("0.25" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.20 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("0.20" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.15 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("0.15" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.10 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% rename("0.10" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
	foreach(m = mm, .combine = bind_rows) %do% {
		s=1
		s = s + 0.05 * m
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("0.05" = `variance.10%`),
		by="mean"
	) %>%
		gather(multiplic_factor, variance, -mean) 

```


```{r, echo=FALSE, eval=FALSE}
my_df %>% ggplot(aes(x=mean, y=variance, color=multiplic_factor)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```


```{r gemme reparam, gamma and negative_binomial, echo=FALSE, eval=FALSE}

## Gamma reparametrised + gamma + negative binomial

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Negative binomial
	foreach(m = mm, .combine = bind_rows) %do% {
			s=3000
			x = rnbinom(10000, mu = m, size = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("negative binomial" = `variance.10%`) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s=10
		s = (s + 0.02 * m) 
		
		v = s^2;
		shape = m^2 / v;
		rate = m / v;
		
		x = rgamma(10000,shape = shape, rate = rate)
			
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("gamma_reparam" = `variance.10%`),
		by="mean"
	) %>% 
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			s=0.4
			x = rgamma(10000,shape = m * s, rate = s)
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		}  %>% rename("gamma" = `variance.10%`),
		by="mean"
	) %>% 
	gather(Label, variance, -mean) 

```


```{r, echo=FALSE, eval=FALSE}
my_df %>% ggplot(aes(x=mean, y=variance, color=Label)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

## Gamma with added variance dependence - slope changing   

**s= 2^(0 + 0* (1/max(mm)) * m)**  
**rgamma(10000,shape =  m * (1/s), rate = (1/s))**  

```{r gamma slope}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 0* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% 
	rename("0" = `variance.10%`) %>% 
		
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 0.5* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("0.5" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 1* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("1" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 1.5* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("1.5" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 2* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("2" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2^(0 + 3* (1/max(mm)) * m)
		x = rgamma(10000,shape =  m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("3" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
		gather(slope, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=slope)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

```{r gamma intercept, echo=FALSE, eval=FALSE}

## Gamma with added variance dependence - intercept changing


mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %do% {
		s= 0 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% 
	rename("0" = `variance.10%`) %>% 
		
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 1 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("0.5" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 2 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("1" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 3 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("1.5" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 4 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("2" = `variance.10%`),
		by="mean"
	) %>%
	
	# Gamma
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
		s= 10 + 2^(1* (1/max(mm)) * m)
		x = rgamma(10000,shape = 1+ m * (1/s), rate = (1/s))
		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	}  %>% rename("3" = `variance.10%`),
		by="mean"
	) %>%
	
	# Poisson
	left_join(
		foreach(m = mm, .combine = bind_rows) %do% {
			
			x = rpois(10000, lambda = m)
			
			c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
		} %>% rename("pois" = `variance.10%`),
		by="mean"
	) %>%
		gather(intercept, variance, -mean) 

```


```{r, echo=FALSE, eval=FALSE}
my_df %>% ggplot(aes(x=mean, y=variance, color=intercept)) + ylab("abs(quantile(x, 0.1) - quantile(x, 0.9))") + geom_line() + theme_bw() 
```

## Negative binomial with added variance dependence - slope changing  

**s= 5000 / 2^(0.0005 * m)**  
**x = rnbinom(10000, mu = m, size =  s )**  

```{r negative binomial slope}

mm = c(seq(2, 10, 1), seq(10, 5000, 10))

foreach(m = mm, .combine = bind_rows) %do% {
		s= 5000 / 2^(0.0005 * m)
		x = rnbinom(10000, mu = m, size =  s )

		c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>% ggplot(aes(x=mean, y=`variance.10%`)) + geom_line() + theme_bw() 
	


```

## Dirichlet


```{r dirichlet}
library(gtools)
mm = c(seq(2, 10, 1), seq(10, 5000, 10))
my_df = 
	
	# Gamma
	foreach(m = mm, .combine = bind_rows) %dopar% {
		
		a = rep(100, 100)
		a[1] = m
		a = a/sum(a)
		s = 1 * 1/min(a)
		
		x = rdirichlet(10000, a * s)[,1]
				c(mean = m, variance = abs(quantile(x, 0.1) - quantile(x, 0.9)))
	} %>%
		gather(intercept, variance, -mean) 

```


```{r}
my_df %>% ggplot(aes(x=mean, y=variance, color=intercept)) + geom_line() + theme_bw() 
```







